---
title: "ANOVA"
author: "Mohd Azmi"
date: "26/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

ANOVA

source: https://www.datanovia.com/en/lessons/anova-in-r/
data source: RM ANOVA DrPH class

other good place to go is https://ademos.people.uic.edu/Chapter21.html

```{r}
library(pacman)
p_load(tidyverse, rstatix, haven)
```

```{r}
confds <- read_sav("Confidence.sav") %>% mutate_if(is.labelled, ~(as_factor(.)))
confds
```


# 1-way ANOVA

group by group

```{r}
confds %>% sample_n_by(group, size = 1)
```

summary stat

```{r}
confds %>%
  group_by(group) %>%
  get_summary_stats(confid1, type = "mean_sd")
```

```{r}
ggplot(confds, aes(x=group, y=confid1)) + geom_boxplot()
```

outlier

```{r}
confds %>% 
  group_by(group) %>% 
  identify_outliers(confid1)
```

normality

```{r}
augment(lm(confid1 ~ group, confds)) %>% 
  ggplot(aes(sample = .resid)) + geom_qq() + geom_qq_line() + facet_grid(.~group)

augment(lm(confid1 ~ group, confds)) %>% 
  shapiro_test(.resid)

confds %>% 
  group_by(group) %>% 
  shapiro_test(confid1)
```

homogeneity of variance

```{r}
augment(lm(confid1 ~ group, confds)) %>%
  ggplot(aes(x=.fitted, y=.resid)) + geom_point() + geom_hline(yintercept = 0)

levene_test(confds, confid1 ~ group)
```

Oneway anova

```{r}
anova_test(confds, confid1 ~ group)
```

```{r}
summary(aov(confid1 ~ group, confds))
```

```{r}
pairwise.t.test(confds$confid1, confds$group, p.adjust.method = "bonf")
```

# 2-way ANOVA

```{r}
confds %>% 
  group_by(group, gender) %>% 
  get_summary_stats(confid1, type = "mean_sd")
```

boxplot

```{r}
ggplot(confds, aes(x=group, y=confid1, colour = gender)) + geom_boxplot()
```

outlier

```{r}
confds %>% 
  group_by(group, gender) %>% 
  identify_outliers(confid1)
```
 

normality

```{r}
augment(lm(confid1 ~ group*gender, confds)) %>%
  ggplot(aes(sample=.resid)) + geom_qq() + geom_qq_line()

augment(lm(confid1 ~ group*gender, confds)) %>%
  shapiro_test(.resid)

confds %>% 
  group_by(group, gender) %>% 
  shapiro_test(confid1)

augment(lm(confid1 ~ group*gender, confds)) %>%
  ggplot(aes(sample=.resid)) + geom_qq() + geom_qq_line() + facet_grid(group ~ gender)
```

Homogeneity of variance

```{r}
levene_test(confds, confid1 ~ group*gender, center = mean)
levene_test(confds, confid1 ~ group*gender, center = median)
```

```{r}
anova_test(confds, confid1 ~ group*gender, detailed = T)
```

```{r}
model.tables(aov(confid1 ~ group*gender, confds))
```

EM Mean

```{r}
confds %>% 
  group_by(gender) %>% 
  emmeans_test(confid1 ~ group, p.adjust.method = "bonferroni") %>% 
  get_emmeans()
```

```{r}
confds %>% 
  filter(gender == "male") %>% 
  emmeans_test(confid1 ~ group, p.adjust.method = "bonferroni") %>% 
  get_emmeans()
```
 
alternative

```{r}
p_load(emmeans)

aovmod <- lm(confid1 ~ group*gender, confds)
summary(aovmod)
```


group
```{r}
emmgroup <- emmeans(aovmod, specs = "group")
emmgroup
pairgroup <- pairs(emmgroup)
pairgroup
```


gender
```{r}
emmgender <- emmeans(aovmod, ~ gender)
emmgender
pairgender <- pairs(emmgender)
pairgender
```


group*gender
```{r}
emmgrgen <- emmeans(aovmod, specs = "group", by = "gender")
emmgrgen
pairgrgen <- pairs(emmgrgen)
pairgrgen
```

group*gender 2?
```{r}
emmgrgen2 <- emmeans(aovmod, specs = "gender", by = "group")
emmgrgen2
pairgrgen2 <- pairs(emmgrgen2)
pairgrgen2
```


group*gender 3?
```{r}
emmgrgen3 <- emmeans(aovmod, ~ group | gender)
emmgrgen3
pairgrgen3 <- pairs(emmgrgen3)
pairgrgen3
```
group*gender 4?
```{r}
emmgrgen4 <- emmeans(aovmod, ~ gender | group)
emmgrgen4
pairgrgen4 <- pairs(emmgrgen4)
pairgrgen4
```

# RM ANOVA

```{r}
confds

confds_long <- confds %>% 
  pivot_longer(c(confid1, confid2, confid3), names_to = "time", values_to = "confidscore") %>% 
  convert_as_factor(id, time)
confds_long
```

summary

```{r}
confds_long %>% 
  group_by(time) %>% 
  get_summary_stats(confidscore, type = "mean_sd")
```

outlier

```{r}
confds_long %>% 
  group_by(time) %>% 
  identify_outliers(confidscore)

```

normality

```{r}
confds_long %>% 
  group_by(time) %>% 
  shapiro_test(confidscore)

```


```{r}
ggplot(confds_long, aes(sample=confidscore)) + geom_qq() + geom_qq_line() + facet_grid(.~time)
```

```{r}
anova_test(confds_long, dv = confidscore, wid = id, within = time)
anova_test(confds_long, dv = confidscore, wid = id, within = time) %>% get_anova_table()

```

pair-wise comparison

```{r}
confds_long %>% 
  pairwise_t_test(confidscore ~ time, paired = T, p.adjust.method = "bonferroni")
```


emmean

```{r}
confds_long %>% 
  emmeans_test(confidscore ~ time, p.adjust.method = "bonferroni") %>% 
  get_emmeans()

confds_long %>% 
  emmeans_test(confidscore ~ time, p.adjust.method = "bonferroni")
```


```{r}
p_load(lme4)

rmanova_mod <- lmer(confidscore ~ time + (1 | id ), confds_long)
```

```{r}
p_load(afex)

rmanova_mod2 <- aov_ez("id", "confidscore", confds_long, within = "time")
rmanova_mod2 # same with anova_test() %>% get_anova_table() above
```

```{r}
emmeans(rmanova_mod2, specs = "time")
```

```{r}
pairs(emmeans(rmanova_mod2, specs = "time"))
```

# Mixed 2-way RM ANOVA

```{r}
confds_long
```


homogeneity of covariance - box m test

```{r}
box_m(confds_long[, "confidscore"], confds_long$group)
```

mixed 2-way RM ANOVA

```{r}
anova_test(confds_long, dv = confidscore, wid = id, between = group, within = time)
anova_test(confds_long, confidscore ~ group + Error(id/time))
get_anova_table(anova_test(confds_long, dv = confidscore, wid = id, between = group, within = time), correction = "GG")
get_anova_table(anova_test(confds_long, dv = confidscore, wid = id, between = group, within = time), correction = "HF")

```


# additional

list of model supported by emmeans http://127.0.0.1:16913/help/library/emmeans/doc/models.html

instead of finding emmean using emmeans package, can also used TukeyHSD

tukeyHSD(model)

```{r}
TukeyHSD(aov(confid1 ~ group, confds))
```

other
